```{r}
library(animation)

createBMLGrid = function(r = c, c = r, ncars = c(red, blue))
  # This function generate a BMLgrid with dimension c*r,
  # with randomly selected red and blue cars.
  # c is the number of columns, r is for row.
  # ncars is the number of cars that may be input as three ways:
  # 1. the number of total cars
  # 2. the number of red cars and blue cars
  # 3. the proportion of cars of all the cells in GMLgrid
  {
  # if input is number of cars, 
  # we consider the number of red and blue cars are equal
  if (length(ncars) == 1)
    ncars = rep(ncars/2, 2)
  # if input is car ratio over total cells of grid
  # ncars = ratio*(total cells) = ratio*r*c
  if (any(ncars < 1))
    ncars[ncars < 1] = ncars[ncars < 1]*r*c
  if ((r|c) <= 0 | sum(ncars) <= 0 |sum(ncars) > r*c) {
    stop('Demensions and ncars should be positive integers， ncars should not be more than product of demension')}
  
  grid = matrix('', r, c)
  # randomly selects cars location out of r*c
  pos = sample(x = r*c, sum(ncars), replace = FALSE)
  # randomly select red and blue cars locations
  grid[pos] = sample(rep(c("red", "blue"), 
                         ceiling(ncars)))[seq_along(pos)]
  # add a class BMLgrid to grid
  class(grid) = c("BMLgrid", class(grid))
  grid
}

Nextlocation = function(g, color, length, ...)
  # This function returns the BMLgrid with the car moves at one step. 
  # g is the BMLgrid, color is the color of car that we assisgned
  # length is either the length of row or column
  # ... is point at either 'row' or 'col' of the current position
  {
  # get the matrix index of cars in certain color, 
  # such as blue and red. 
  current = which(g == color , arr.ind = TRUE)
  # the car moves along its row or col. 
  move = (as.numeric(current[, ...]) + 1)
  # If the car moves to the edge of the grid, 
  # then it starts from the begining: 
  # the first postions of its current row or column
  move[move > length] = 1
  nextl = current
  # only the row or column changed compare to the car's current location
  nextl[, ...] = move
  # If the next position that the car tends to move is not empty,
  # they it is stucked and stay at its current postion. 
  check = (g[nextl] != '')
  nextl[check, ] = current[check, ]
  g[current] = ''
  g[nextl] = color
  g
}


moveCars = function(g, color)
  # This function return the BMLgrid 
  # with either red or blue car moves by 1 step
  # color means the color of car that moves
  {
  if(color == 'blue')
    # The blue car moves up along its column,
    # so the column of its position stays the same and
    # every move increase the row at 1 unit. 
    # As long as it gets the edge of the grid
    # (the end of the number of row), 
    # it starts at the 1st row of its positon
    Nextlocation(g = g, color, length = nrow(g), 'row')
    else
    # The red car moves right along its row,
    # so the row of its position stays the same and
    # every move increase the column at 1 unit. 
    # As long as it gets the end of the number of column, 
    # it starts at the 1st column of its positon.
    Nextlocation(g = g, color, length = ncol(g), 'col')
}



runBMLGrid = function(g, numsteps)
  # This function return the BMLgrid 
  # after the cars moves by a certain number of steps.
  # g is the first stage of BMLgrid
  # numsteps is the number of moving steps. 
  {
  for (i in (1:numsteps)){
    # if the time periods is odd, then the blue car moves, 
    # otherwise, the red car moves. 
    if (i%%2 == 1) 
      g = moveCars(g, color = 'blue')
    else
      g = moveCars(g, color = 'red')
    }
  g
}
```
Plot
```{r}
plot.BMLgrid = function(g)
  # This function displays the grid
  {
  # Match the color to generate a matrix with numbers not characters
  # each number represents one color
  colmat = match(g, c('red', 'blue', ''))
  g = matrix(colmat, nrow(g), ncol(g))
  image(t(g), col = c('red', 'blue', 'white'), axes = FALSE, main = 'Car Move Simulation')
  box()
  }
```
Summary
```{r}
summary.BMLgrid = function(g) 
  # This summary funtion returns general inforamtion of the BMLgrid:
  # the dimension, the number of cars, red cars and blue cars, 
  # and the ratio of cars, blue cars, red cars over total cells of BMLgrid
  {
  NUM = rbind(Totcar = sum(g != ''), 
                       redcar = sum(g == 'red'), 
                       bluecar = sum(g == 'blue'))
  colnames(NUM) = 'NUM'
  RATIO = NUM/prod(dim(g))
  colnames(RATIO) = 'RATIO'
  list(Demention = dim(g),
       CarsNum = cbind(NUM, RATIO)
       )
}
```
The number of car that moved, that were bloked and average velocity.
```{r}
NumCarMBV = function(oldpos, newpos)
  # This function returns a list of the number of cars that moved, 
  # were blocked and the velocity at one time period.
  # oldpos is the postion at time period t
  # newpos is the postion at time period t + 1
  {
  # unmatched cell is the moved car and its next empty cell
  # so the number of move car is total unmached cell divided by 2. 
  redmove = sum(!(oldpos == 'red') == (newpos == 'red'))/2
  bluemove = sum(!(oldpos == 'blue') == (newpos == 'blue'))/2
  # find which color of cars move at this time
  move = redmove + bluemove
  summary = summary.BMLgrid(oldpos)$CarsNum[, 'NUM']
  totalcar = summary['Totcar']
  # The number of stuck car is total cars minus move cars
  carblock = totalcar - carmove
  # We define the Velocity as ratio of the move car to the total car. 
  # Average velocity is the average of blue and red car velocity. 
  velocity = mean(c(redmove/summary['redcar'], bluemove/summary['bluecar']))
  list(NumCarMove = carmove, 
       NumCarBlock = carblock, 
       velocity = velocity)
}

Allmbv = function(g.out)
  # This function returna a list of the number of cars 
  # that moved, were blocked and the velocity at every time period.
  # g.out is a list of every stage BMLGrid 
  # during a certain number of steps. 
  {
  all = c()
  for (i in (2:length(g.out))){
    oldpos = g.out[[i-1]]
    newpos = g.out[[i]]
    all[[i]] = NumCarMBV(oldpos, newpos)
}
all
}

getall

g = createBMLGrid(10, 10, 0.5)
g.all = AllstageBMLgrid(g, numsteps = 100)
saveGIF(sapply(g.all, plot.BMLgrid))
system.time(runBMLGrid(g, numsteps = 10000))
```
The performance of the code
```{r}
library(ggplot2)
library(reshape)
time = function(g, numsteps){
  t = system.time(runBMLGrid(g, numsteps))['elapsed']
}
# different grid sizes with numsteps 1000 and rho 0.5
alltime = function(...){
  g = sapply(seq(10, 1010, by = 100), createBMLGrid, ...)
  sapply(g, time, numsteps = 200)
}
t0 = alltime(ncars = 0.1)
t1 = alltime(ncars = 0.3)
t2 = alltime(ncars = 0.5)
t3 = alltime(ncars = 0.7)
t4 = alltime(ncars = 0.9)
time = as.data.frame(cbind(t0, t1, t2, t3, t4))
rownames(time) = seq(10, 1010, by = 100)
colnames(time) = paste0(rep('rho', 5), seq(0.1, 0.9, by = 0.2))
time = cbind(time, dim = seq(10, 1010, by = 100)^2)
par(mfrow = c(1, 1))
time = melt(time, id = 'dim', variable_name = 'difdim')
ggplot(time, aes(dim, value, col = factor(difdim),linetype = factor(difdim), shape = factor(difdim))) + geom_line(size = 2) + labs(title = 'The Elapsed Time, Size and Density in R', x = 'Number of Grid Cells', y = 'Elapsed Time/s') + geom_point(size = 4) + theme(text = element_text(size=20), axis.text=element_text(colour="black", size = 20)) 
```
Analysis
```{r}
getallvel = function(g, numsteps)
  #This function returns a list of average velocity at every time period
  {
  g.all = AllstageBMLgrid(g, numsteps)
  unlist(sapply(Allmbv(g.all), '[', 3))
}

multiplot = function(vel, ...)
  # This function returns multiple plots
  # vel a multiple list of average velocity 
  # ... some format of plot such as lab, main. 
  {
  sapply(vel, plot, type = 'l', mex = 5, lwd = 5, ...)
}
```
```{r}
# 10*10
g = lapply((1:6)*0.1, createBMLGrid, c = 10, r = 10)
vel = lapply(g, getallvel, numsteps = 100)
p = multiplot(vel, ylab = 'Velocity', main = 'Velocity From Time 10 to 45(dim = 10*10)', xlab = 'Number of Steps', xlim = c(10, 45))

# 100*100
g = lapply((2:7)*0.1, createBMLGrid, c = 100, r = 100)
vel = lapply(g, getallvel, numsteps = 3000)
par(mfrow = c(3, 2))
multiplot(vel, ylab = 'Velocity', main = 'Velocity From Time 10 to 45(dim = 10*10)', xlab = 'Number of Steps', xlim = c(0, 1000))

g = lapply((35:40)*0.01, createBMLGrid, c = 100, r = 100)
vel = lapply(g, getallvel, numsteps = 10000)
par(mfrow = c(3, 2))
sapply(vel, plot, type = 'l', ylab = 'Velocity', xlab = 'Number of Time Periods', xlim = c(400, 500), main = 'Velocity From Time 400 to 500', xlim = c())

# 500*500
g = lapply((3:7)*0.1, createBMLGrid, c = 500, r = 500)
vel = lapply(g, getallvel, numsteps = 200)
par(mfrow = c(3, 2))
sapply(vel, plot, type = 'l', ylab = 'Velocity', main = 'Velocity From Time 40 to 60')
```

```{r}
# 100*100
g = lapply((3:7)*0.1, createBMLGrid, c = 100, r = 100)
lapply(g, summary)
vel = lapply(g, getallvel, numsteps = 200)
par(mfrow = c(3, 2))
sapply(vel, plot, type = 'l', ylab = 'Velocity', main = 'Velocity From Time 40 to 60')
````


慢的 版本2
```{r}
moveCars = function(g, color){
  white = which(g == '', arr.ind = TRUE)
  if(color = 'blue'){
    blue = which(g == color , arr.ind = TRUE)
    row = (as.numeric(blue[,'row']) + 1) %% nrow(g)
    row[row == 0] = nrow(g)
    nextblue = cbind(row, col = blue[, 'col'])
    nextblue[g[nextblue] != '', ] = blue[g[nextblue] != '', ]
    g[blue] = ''
    g[nextblue] = 'blue'
  }else if (color == 'red'){
    red = which(g == color, arr.ind = TRUE)
    col = (as.numeric(red[,'col'])+1) %% ncol(g)
    col[col == 0] = ncol(g)
    nextred = red
    nextred[, 'col'] = col
    nextred[g[nextred] != '',] = red[g[nextred] != '', ]
    g[red] = ''
    g[nextred] = 'red'
  }
  g
}
```
慢的 版本1
```{r}
car_next_location = function(currentpos, nextpos, white){ 
  # check if next position is empty
  checkempty = match(interaction(white[, 'row'], white[, 'col']), 
                     interaction(nextpos[, 'row'], nextpos[, 'col'])
  )
  if (any(is.na(checkempty))){
    if (all(is.na(checkempty)))
      nextpos = currentpos else{
        checkempty = checkempty[!is.na(checkempty)]
        nextpos[-checkempty, ] = currentpos[-checkempty, ]
      }}
  nextpos
}

moveCarsold = function(g, col = ...){
  white = which(g == '', arr.ind = TRUE)
  if(col == 'blue'){
    blue = which(g == col, arr.ind = TRUE)
    row = (as.numeric(blue[,'row']) + 1) %% nrow(g)
    row[row == 0] = nrow(g)
    nextblue = cbind(row, col = blue[, 'col'])
    nextblue = car_next_location(blue, nextblue, white)
    g[blue] = ''
    g[nextblue] = 'blue'
  }else if (col == 'red'){
    red = which(g == col, arr.ind = TRUE)
    col = (as.numeric(red[,'col'])+1)%%ncol(g)
    col[col == 0] = ncol(g)
    nextred = cbind(row = red[, 'row'], col = col)
    nextred = car_next_location(red, nextred, white)
    g[red] = ''
    g[nextred] = 'red'
  }
  g
}


NumCarMove = function(oldpos, newpos)
  # This function returns the number of cars that can move。
  # oldpos is the postion at time period t
  # newpos is the postion at time period t + 1
  {
  # unmatched cell is the moved car and its next empty cell
  # so the number of move car is total unmached cell divided by 2. 
  sum(!oldpos == newpos)/2
}

NumCarStuck = function(oldpos, newpos)
  #This function returns the number of cars that gets stucked
  {
  # The number of stuck car is total cars minus move cars
  sum(oldpos != '') - sum(!oldpos == newpos)/2
}

Velocity = function(oldpos, newpos)
  # We define the Velocity as ratio of the move car to the total car. 
  # This function return velocity. 
  {
  (sum(!oldpos == newpos)/2)/(sum(oldpos != ''))
}



Allmsv = function(g.out)
  # 
  {
  move = c()
  stuck = c()
  v = c()
  for (i in (2:length(g.out))){
    oldpos = g.out[[i-1]]
    newpos = g.out[[i]]
    move[i] = NumCarMove(oldpos, newpos)
    stuck[i] = NumCarStuck(oldpos, newpos)
    v[i] = Velocity(oldpos, newpos)
}
list(carmove=move, carstuck = stuck, velocity = v)
}
